---
title: "Practical Machine Learning - Project"
author: "Oscar Peña del Rio"
date: "27/12/2015"
output: html_document
---

First, we load all the libraries we will be using within our analysis, and load the data itself (we suppose the data is present in the same folder as the R script).

```{r}
library(caret)
library(randomForest)
library(rattle)
library(rpart.plot)
library(RColorBrewer)

training_data <- read.csv('pml-training.csv')
testing_data <- read.csv('pml-testing.csv')
```

Print how many observations and variables we have in each dataset.

```{r}
dim(training_data)
dim(testing_data)
```

In order to remove columns where all observations are NA, we perform the following processing step.

```{r}
training_data <- training_data[, colSums(is.na(training_data)) == 0]
testing_data <- testing_data[, colSums(is.na(testing_data)) == 0]
```

We still have some columns we are not interested in, so we are going to remove some of them manually from the *testing_data* set, and the drop all the columns from *training_data* which are not present in *testing_data* (they are NOT sensor related). We will save the *classe* column and append it afterwards to the *training_data* set.

```{r}
testing_data <- subset(testing_data, select=-c(
  X, raw_timestamp_part_1, raw_timestamp_part_2, 
  cvtd_timestamp, new_window, num_window, problem_id
))

dim(testing_data)

col_names_to_keep <- colnames(testing_data)
classe <- training_data$classe

training_data <- training_data[, (colnames(training_data) %in% col_names_to_keep)]
training_data$classe <- classe
dim(training_data)
```

Let's begin our prediction!

First, we create a partition on our *training_data* for cross validation, 70% will form the *train_data* set and the remaining 30% the *test_data* set.

```{r}
# set seed to today's date
set.seed(20151227)

for_training <- createDataPartition(training_data$classe, p=0.7, list=FALSE)

train_data <- training_data[for_training,]
test_data <- training_data[-for_training,]
```

For the prediction model, we will apply the Random Forest algorithm. First, we set the control parameters for the *train()* function, using 10-folds for cross validation.

```{r}
RF_model <- randomForest(
  train_data$classe ~ .,
  data=train_data,
  ntree=250
)

RF_model
```

Now, we build our prediction tool, and check it against the *test_data* set.

```{r}
RF_prediction <- predict(RF_model, test_data)

confusion_matrix <- confusionMatrix(test_data$classe, RF_prediction)

model_accuracy <- postResample(RF_prediction, test_data$classe)
model_error <- 1 - as.numeric(confusion_matrix$overall[1])
```

Finally, we print all the results:

```{r}
confusion_matrix
model_accuracy
model_error
```

And the prettyfied decision tree of the *randomForest()* function:

```{r}
decision_tree <- rpart(
  train_data$classe ~ .,
  data=train_data,
  method='class'
)

fancyRpartPlot(decision_tree)
```

## Conclusions

We have an accuracy of 99.39%, with an error rate of 0.61%. They are quite good results for a *quick* analysis. So we launch it against the initial *testing_data* set, in order to predict their outcome values.

```{r}
prediction <- predict(RF_model, testing_data)

prediction
```